---
title: "OWID_up_down"
author: "Matthew Morriss"
date: '2022-04-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(owidR)
library(here)
library(lubridate)
library(reshape2)
library(ggraph)
library(tidygraph)
library(igraph)
library(showtext)
library(ggtext)
library(cowplot)

```

## download data

This is a markdown document to download water qualtiy data from OWID and look to make a curvy arc plot of it URBan vs rural

```{r cars}
#determine what data to download

basic_water <-  as.data.frame(owid_search("water"))

dw_df <- owid(chart_id = basic_water$chart_id[56])

```

## Clean up OWID df

These data are pretty messy. Need to do some significant cleaning to see:
1) what years have the most data for countries
2) remove non-nonsensical years
3) look at last 10 years of data
```{r clean data, echo=FALSE}
cyr <- 2022

cdw_df <- dw_df %>%
  filter(year > (cyr - 10))


#look at how many years have good data for drinking water in the rural areas
cdw_df %>%
  group_by(year) %>%
  summarise(unique = n_distinct(`Drinking water - At least basic - Rural - Percent`))


#look at how many years have good data for drinking water in the urban areas

cdw_df %>%
  group_by(year) %>%
  summarise(unique = n_distinct(`Drinking water - At least basic - Urban - Percent`))

```

Based on my simple analysis above, we want to examine data in 2013 as there are ~155 data points for different countries Urban vs Rural. 

```{r,results='hide', echo=FALSE}
#filter to year 2013, remove NA rows 
cdw_df <- cdw_df %>%
  filter(., year == 2013) %>%
  drop_na(`Drinking water - At least basic - Rural - Percent`, `Drinking water - At least basic - Urban - Percent`)


cdw_df <- cdw_df %>%
  select(entity,
         year,
         `Drinking water - At least basic - Rural - Percent`,
         `Drinking water - At least basic - Urban - Percent`) %>%
  reshape2::melt(id = c("entity","year"))



```
Now I have a clean dataset for ~173 entities drinking water access in urban/rural areas for 2013. Time to start looking into the up-down curve plot.

Following along with the coding example below. Want to find country with largest difference between urban and Rural
```{r, echo=FALSE}
cdw_df %>%
  group_by(entity, year) %>%
  arrange(entity, year) %>%
  mutate(diff = value - lag(value)) %>%
  ungroup() %>%
  # filter(., entity == "United States")
  # slice_max(diff, n = 1)
  filter(.,entity == "Eritrea")


cdw_df %>%
  group_by(entity, year) %>%
  arrange(entity, year) %>%
  mutate(diff = value - lag(value)) %>%
  ungroup() %>%
  # filter(., entity == "United States")
  # slice_max(diff, n = 1)
  filter(.,entity == "United States")


```
Eritrea has the largest difference between rural and urban areas in terms of basic water access.

# Plotting
My next steps will be to separate out the Urban and rural and start to build my arc diagrams. 

```{r urban, echo=FALSE}
urban_df <- cdw_df %>%
  dplyr::filter(., grepl("Urban", variable)) %>%
  mutate(value = round(value))

urban_df %>%
  summarize(ncountries = n_distinct(entity),
            max_urban = max(value))

links_urban <- data.frame(source = 0,
                         value = c(1:100))


links <- links_urban %>%
  left_join(urban_df, by = "value") %>%
  mutate(country = if_else(is.na(entity),"",entity),
         group = factor(case_when(is.na(entity) ~ 1,
                                  value == 66 ~3,
                                  # value == 100 ~ 4,
                                  TRUE ~ 2))) %>%
  select(source,
         destination = value, 
         group)%>%
  group_by(destination) %>%
  mutate(width = n())
  

# graph <- graph_from_data_frame(links)



g <- links %>%
  graph_from_data_frame() %>%
  ggraph(layout="linear") + #call the ggpraph function 
  geom_edge_arc(aes(edge_colour = group,  #three color groups, no data, data, and value highlighted
                    edge_width = width,   # line width set by number of countries with similar ages
                    edge_alpha = group)) + # transparency by group number...
  geom_node_text(aes(label = name),
                 repel = FALSE,
                 size = 6,
                 nudge_y = -1,
                 color = rep(c("white", NA, NA, NA, NA, NA),
                 length.out = 101))+ #add labels to the nodes, from "names" which idk where that came from in the function calls previously
  scale_edge_color_manual(values = c("black","lavender", "darkgoldenrod1")) +  #adding gray to major lines, staying green for highlight
  scale_edge_alpha_manual(values = c(0, .15, 1)) + #set group 3 to only one not transparent
  labs(title = "Basic Safe Drinking Water",
       subtitle = "<span style = 'color: white;'>The Rate of access to at least safe drinking water in</span> <span style = 'color: darkgoldenrod1;'>Urban</span> <span style = 'color: white;'>and</span> <span style = 'color: darkorchid;'>Rural</span> <span style = 'color: white;'>areas</span>"
       ) +
  theme_void() +
  theme(
    # plot.margin = margin(t = 40, r = 50, b = 0, l = 50),
        panel.background = element_rect(fill = "black", color = "black"),
        plot.background = element_rect(fill = "black", color = "black"),
        legend.position = "none",
        # plot.title = element_markdown(size = 96),
        plot.title = element_text(colour = "white"),
        # plot.subtitle = element_text(color = "white"),
        plot.subtitle = element_markdown())

g


ggsave(here("Figures","Urban.pdf"), width = 16.5, height = 11.7, units = "in", dpi = 320)
```


Now plot the rural access to clean drinking water in eritrea
```{r rural, echo=FALSE}
rural_df <- cdw_df %>%
  dplyr::filter(., grepl("Rural", variable)) %>%
  mutate(value = round(value))

rural_df %>%
  summarize(ncountries = n_distinct(entity),
            max_rural = max(value))

links_rural <- data.frame(source = 0,
                         value = c(1:100))


links <- links_rural %>%
  left_join(rural_df, by = "value") %>%
  mutate(country = if_else(is.na(entity),"",entity),
         group = factor(case_when(is.na(entity) ~ 1,
                                  value == 6 ~3,
                                  TRUE ~ 2))) %>%
  select(source,
         destination = value, 
         group)%>%
  group_by(destination) %>%
  mutate(width = n())
  

graph <- graph_from_data_frame(links)



g2 <- ggraph(graph, layout="linear") + #call the ggpraph function 
  geom_edge_arc(aes(edge_colour = group,  #three color groups, no data, data, and value highlighted
                    edge_width = width,   # line width set by number of countries with similar ages
                    edge_alpha = group)) + # transparency by group number...
 geom_node_text(aes(label = name),
                 repel = FALSE,
                 size = 6,
                 nudge_y = -1,
                 color = rep(c("white", NA, NA, NA, NA, NA),
                 length.out = 101))+
  scale_edge_color_manual(values = c("black","lavender", "darkorchid")) +  #adding gray to major lines, staying green for highlight
  scale_edge_alpha_manual(values = c(0, .15, 1)) + #set group 3 to only one not transparent
  # labs(title = "Rate of access to at least basic safe drinking water in Rural areas") +
  theme_void() +
  theme(
    # plot.margin = margin(t = 40, r = 50, b = 0, l = 50),
        panel.background = element_rect(fill = "black", color = "black"),
        plot.background = element_rect(fill = "black", color = "black"),
        legend.position = "none")
        # plot.title = element_text(colour = "#7F3D17"))+
  # scale_y_continuous(trans = "reverse")+
  # scale_y_reverse()

g2

ggsave(here("Figures","Rural.pdf"), width = 16.5, height = 11.7, units = "in", dpi = 320)
```


Now I want to attempt to stitch the two plots together
```{r}
#setup position for frame 
cowplot::ggdraw(xlim = c(0,0.97),
                ylim = c(-1,1.25))+
  #draw first plot in top of frame
  cowplot::draw_plot(g,
                     x = 0,
                     y = 0)+
  # draw second plot in bottom half of frame
  cowplot::draw_plot(g2,
                     x = 0,
                     y = -1)


```


```{r}

set_graph_style(plot_margin = margin(1,1,1,1))
graph <- as_tbl_graph(highschool)


ggraph(graph, layout = 'linear') + 
  geom_edge_arc(aes(colour = factor(year)),
                fold = FALSE)
```