---
title: "OWID_up_down"
author: "Matthew Morriss"
date: '2022-04-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(owidR)
library(here)
library(lubridate)
library(reshape2)
library(ggraph)
library(tidygraph)
library(igraph)
library(showtext)
library(ggtext)
library(cowplot)

```

## download data

This is a markdown document to download water quality data from OWID and look to make a curvy arc plot of it - urban vs rural

```{r cars}
#determine what data to download

basic_water <-  as.data.frame(owid_search("water"))

dw_df <- owid(chart_id = basic_water$chart_id[56])

```

## Clean up OWID df

These data are pretty messy. Need to do some significant cleaning to see:
1) what years have the most data for countries
2) remove nonsensical years
3) look at last 10 years of data
```{r clean data, echo=FALSE}
cyr <- 2022

cdw_df <- dw_df %>%
  filter(year > (cyr - 10))

#look at how many years have good data for drinking water in the rural areas
cdw_df %>%
  group_by(year) %>%
  summarise(unique = n_distinct(`Drinking water - At least basic - Rural - Percent`))


#look at how many years have good data for drinking water in the urban areas
cdw_df %>%
  group_by(year) %>%
  summarise(unique = n_distinct(`Drinking water - At least basic - Urban - Percent`))

```

Based on my simple analysis above, we want to examine data in 2013 as there are ~155 data points for different countries Urban vs Rural. 

```{r,results='hide', echo=FALSE}
#filter to year 2013, remove NA rows 
cdw_df <- cdw_df %>%
  filter(., year == 2013) %>%
  drop_na(`Drinking water - At least basic - Rural - Percent`, `Drinking water - At least basic - Urban - Percent`)

cdw_df <- cdw_df %>%
  select(entity,
         year,
         `Drinking water - At least basic - Rural - Percent`,
         `Drinking water - At least basic - Urban - Percent`) %>%
  reshape2::melt(id = c("entity","year"))

```
Now I have a clean dataset for ~173 entities drinking water access in urban/rural areas for 2013. Time to start looking into the up-down curve plot.

Following along with the coding example below. Want to find global average difference between urban and Rural
```{r, echo=FALSE}
cdw_df %>%
  group_by(entity, year) %>%
  arrange(entity, year) %>%
  mutate(diff = value - lag(value)) %>%
  ungroup() %>%
  filter(.,entity == "World")

```
The global difference between rural and urban areas (in terms of basic water access) in 2013 was 16.5%.

# Plotting
My next steps will be to separate out the Urban and rural and start to build my arc diagrams. 

```{r urban, echo=FALSE}
library(showtext)

# add font
font_fam = "Source Sans Pro"
font_add_google(font_fam, regular.wt = 300, bold.wt = 700) 
showtext_opts(dpi = 300)
showtext_auto(enable = TRUE)

# common theme for both plots
theme_arc <- function(){
  theme_void(base_size = 32) +
  theme(
      text = element_text(family = font_fam, color = "white"),
        panel.background = element_rect(fill = "black", color = "black"),
        plot.background = element_rect(fill = "black", color = "black"),
        legend.position = "none",
        plot.title = element_text(colour = "white", family = font_fam, face = "bold"),
        plot.subtitle = element_text(color = "white", size=30))
}

```

```{r}

# Prep data for plotting
urban_df <- cdw_df %>%
  dplyr::filter(., grepl("Urban", variable)) %>%
  mutate(value = round(value))

# create link data for graph
links_urban <- data.frame(source = 0,
                         value = c(1:100)) %>%
  left_join(urban_df, by = "value") %>%
  mutate(country = if_else(is.na(entity),"",entity),
         group = factor(case_when(is.na(entity) ~ 1,
                                  entity == 'World' ~ 3,
                                  TRUE ~ 2))) %>%
  select(source,
         destination = value, 
         group)%>%
  group_by(destination) %>%
  mutate(width = n())

# plot urban graph
g <- links_urban %>%
  graph_from_data_frame() %>%
  ggraph(layout="linear") + #call the ggpraph function 
  geom_edge_arc(aes(edge_colour = group,  #three color groups, no data, data, and value highlighted
                    edge_width = width,   # line width set by number of countries with similar ages
                    edge_alpha = group # transparency by group number...
                    )
                ) + 
  geom_node_text(aes(label = name),
                 repel = FALSE,
                 size = 6,
                 nudge_y = -1,
                 color = rep(c("white", NA, NA, NA, NA, NA, NA, NA, NA, NA),
                 length.out = 101))+ #add labels to the nodes, from "names" which idk where that came from in the function calls previously
  scale_edge_color_manual(values = c("black","darkgoldenrod1", "white")) +  
  scale_edge_alpha_manual(values = c(0, 1, 1)) + #set groups 2 and 3 to not transparent, set group 1 (entity==NA) to fully transparent
  labs(title = "Basic access to safe drinking water",
       subtitle = "Comparing access to safe drinking water in urban and rural areas around the world") +
  theme_arc()
g

ggsave("Figures/Urban.png", width = 16, height = 9, units = "in", dpi = 300)
```


Now plot the rural access to clean drinking water in eritrea
```{r rural, echo=FALSE}
# prep graph data
rural_df <- cdw_df %>%
  dplyr::filter(., grepl("Rural", variable)) %>%
  mutate(value = round(value))

links_rural <- data.frame(source = 0,
                         value = c(1:100)) %>%
  left_join(rural_df, by = "value") %>%
  mutate(country = if_else(is.na(entity),"",entity),
         group = factor(case_when(is.na(entity) ~ 1,
                                  entity == 'World' ~ 3,
                                  TRUE ~ 2))) %>%
  select(source,
         destination = value, 
         group)%>%
  group_by(destination) %>%
  mutate(width = n())

# make rural graph
g2 <- links_rural %>% 
  graph_from_data_frame() %>%
  ggraph(layout="linear") + #call the ggpraph function 
  geom_edge_arc(aes(edge_colour = group,  #three color groups, no data, data, and value highlighted
                    edge_width = width,   # line width set by number of countries with similar ages
                    edge_alpha = group),
                strength = -1) + # transparency by group number...
 geom_node_text(aes(label = name),
                 repel = FALSE,
                 size = 0,
                 nudge_y = -1,
                 color = rep(c("white", NA, NA, NA, NA, NA, NA, NA, NA, NA),
                 length.out = 101))+
  scale_edge_color_manual(values = c("black","darkorchid", "white")) +  
  scale_edge_alpha_manual(values = c(0, 1, 1)) + #set groups 2 and 3 to not transparent, set group 1 (entity==NA) to fully transparent
  theme_arc()

g2

ggsave("Figures/Rural.png", width = 16, height = 8, units = "in", dpi = 300)
```


Stitch the two plots together
```{r}

plot_margin <- 0.025

# black bar for bottom of plot
box_black <- grid::rectGrob(
  x = 0, y = 0, 
  width = 16, height = 2,
  gp = grid::gpar(fill = "black", alpha = 1, col = 'black')
)
canvas <- grid::rectGrob(
  x = 0, y = 0, 
  width = 16, height = 9,
  gp = grid::gpar(fill = "black", alpha = 1, col = 'black')
)

#setup position for frame 
cowplot::ggdraw(xlim = c(0,1),
                ylim = c(0.1,1))+
  # black background
  draw_grob(canvas,
            x = 0, y = 1,
            height = 9, width = 16,
            hjust = 0, vjust = 1)+
  # draw second plot in bottom half of frame
  cowplot::draw_plot(g2,
                     x = 0,
                     y = 0.125,
                     height = 0.45)+
  #draw first plot in top of frame
  cowplot::draw_plot(g,
                     x = 0,
                     y = 0.5,
                     height = 0.5) +
  draw_grob(box_black,
            x = 0, y = 0.1,
            height = 1, width = 16,
            hjust = 0, vjust = 1)+
  # creator
  draw_label("Matthew Morriss, USGS\nData: OWID Water, Sanitation, and Hygiene Data Explorer,WHO/UNICEF Joint Monitoring Program for Water Supply and Sanitation", 
             x = 1-plot_margin, 
             y = 0.12, 
             fontface = "italic", 
             color = "white",
             size = 16, 
             hjust = 1, vjust = 0,
             fontfamily = font_fam,
             lineheight = 1.1) 

  
ggsave("Figures/owid_upddown.png", width = 14, height = 16, units = "in", dpi = 300)
```
