---
title: "GNP_figures"
author: "Matthew Morriss"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggpubr)
library(reshape2)
library(stringr)
library(here)
library(CGPfunctions)
library(ggrepel)
# rm(list = ls())
```

## load data

```{r}
# g_df_2015 <- read.csv("D:\\OneDrive - DOI\\Rock_Glaciers\\Proposals\\NPS_proposal_2021\\tables\\GNP_glaciers_2015.csv")
# g_df_2005 <- read.csv("D:\\OneDrive - DOI\\Rock_Glaciers\\Proposals\\NPS_proposal_2021\\tables\\GNP_glaciers_2005.csv")
# g_df_1998 <- read.csv("D:\\OneDrive - DOI\\Rock_Glaciers\\Proposals\\NPS_proposal_2021\\tables\\GNP_Glaciers_1998.csv")
# g_df_1966 <- read.csv("D:\\OneDrive - DOI\\Rock_Glaciers\\Proposals\\NPS_proposal_2021\\tables\\GNP_glaciers_1966.csv")
# 
# g_df <- left_join(g_df_2015, g_df_2005, by  = "GLACNAME") %>%
#   left_join(.,g_df_1998, by = "GLACNAME")%>%
#   left_join(.,g_df_1966, by = "GLACNAME")
# 
# g_df <- g_df %>%
#   select(GLACNAME, Area2015, Area2005, Area1998, Area1966)
# 
# g_df <- g_df %>%
#   melt()
# 
# colnames(g_df) <- c("Glacier Name",
#                     "Year",
#                     "Area m^2")
# 
# regexp <- "[[:digit:]]+"
# g_df$Year <- str_extract(g_df$Year, regexp)
# 
# 
# write.csv(g_df,paste(here("Data"),"GNP_gdata.csv"))
```

## Plotting GNP glaciers
```{r pressure, echo=FALSE}
n_glac <- 15

g_df <- read.csv(here("Data","Data GNP_gdata.csv"))
g_df <- g_df %>%
  mutate(Area.sq.km = round(Area.m.2/1000/1000, digits = 2))



big_g <- g_df %>%
  filter(., Year == 1966) %>%
  top_n(n_glac, Area.sq.km) %>%
  # sample_n(.,10) %>%
  select(Glacier.Name)

g_df <- g_df %>%
  filter(., Glacier.Name %in% big_g$Glacier.Name)

g_df$Year <- as.factor(g_df$Year)



newggslopegraph(g_df,
                Year,
                Area.sq.km,
                Glacier.Name,
                Title = "10 Largest Glaciers in Glacier National Park",
                SubTitle = bquote("Area in" ~km^2),
                Caption = NULL,
                DataTextSize = 2.5,
                DataLabelPadding = 0.05,
                WiderLabels = TRUE)
  

```

What if I instead wanted to look at a normalized plot of percent change through time, bringing these glaciers to a single point at the beginning of the time series domain.

```{r}
g_df <- read.csv(here("Data","Data GNP_gdata.csv"))
g_df <- g_df %>%
  mutate(Area.sq.km = round(Area.m.2/1000/1000, digits = 2))


big_g <- g_df %>%
  filter(., Year == 1966) %>%
  top_n(10, Area.sq.km) %>%
  # sample_n(.,15) %>%
  select(Glacier.Name)

# small_g <- g_df %>%
#   filter(., Year == 1966) %>%
#   top_n(-5, Area.sq.km) %>%
#   select(Glacier.Name)
# 
# big_g <- rbind(big_g, small_g)

g_df <- g_df %>%
  filter(., Glacier.Name %in% big_g$Glacier.Name) %>%
  group_by(Glacier.Name) %>%
  arrange(Year) %>%
  mutate(lag = lag(Area.sq.km)) %>%
  mutate(pct.change = (100*(Area.sq.km - lag)/lag))%>%
  mutate(cum_sum_pct_chg = round(cumsum(if_else(is.na(pct.change),0,pct.change))+ pct.change*0), digits = 2)



g_df[is.na(g_df)] <- 0
  

g_df$Year <- as.factor(g_df$Year)



newggslopegraph(g_df,
                Year,
                cum_sum_pct_chg,
                Glacier.Name,
                Title = "10 Largest Glaciers in Glacier National Park",
                SubTitle = "Cumulative Percent Loss of Area",
                Caption = "Data: Fagre et al., 2017",
                DataTextSize = 2.5,
                DataLabelPadding = 0.05,
                WiderLabels = TRUE)


```

Next up is trying to make the slope plot in GGplot 


```{r, echo=FALSE}
g_df <- read.csv(here("Data","Data GNP_gdata.csv"))
g_df <- g_df %>%
  mutate(Area.sq.km = round(Area.m.2/1000/1000, digits = 2))


big_g <- g_df %>%
  filter(., Year == 1966) %>%
  top_n(10, Area.sq.km) %>%
  # sample_n(.,15) %>%
  select(Glacier.Name)


g_df <- g_df %>%
  filter(., Glacier.Name %in% big_g$Glacier.Name) %>%
  group_by(Glacier.Name) %>%
  arrange(Year) %>%
  mutate(lag = lag(Area.sq.km)) %>%
  mutate(pct.change = (100*(Area.sq.km - lag)/lag))%>%
  mutate(cum_sum_pct_chg = round(cumsum(if_else(is.na(pct.change),0,pct.change))+ pct.change*0), digits = 2)



g_df[is.na(g_df)] <- 0

```

Plot below with lines colored by glacier
!!!! This is the plot used in the final post !!!!!
```{r}
######
MySpecial <- list(
  
  scale_x_continuous(position = "top",
                   breaks = c(1966, 1998, 2005, 2015),
                   # expand = expansion(mult = c(0.01, 0.285)),
                   limits =  c(1966, 2030)),
  theme_bw(),
  # Format tweaks
  # Remove the legend
  theme(legend.position = "none"),
  # Remove the panel border
  theme(panel.border     = element_blank()),
  
  # Remove just about everything from the y axis
  theme(axis.title.y     = element_blank()),
  theme(axis.text.y      = element_blank()),
  theme(panel.grid.major.y = element_blank()),
  theme(panel.grid.minor.y = element_blank()),
  # Remove a few things from the x axis and increase font size
  theme(axis.title.x     = element_blank()),
  theme(panel.grid.major.x = element_blank()),
  theme(panel.grid.minor.x = element_blank()),
  theme(axis.text.x.top      = element_text(size=12)),
  # Remove x & y tick marks
  theme(axis.ticks       = element_blank()),
    # Format title & subtitle
  theme(plot.title       = element_text(size=14, face = "bold", hjust = 0.5)),
  theme(plot.subtitle    = element_text(hjust = 0.5))
  # theme(plot.margin = unit(c(0.1, 5, 0.1, 0.1), "cm"))
)

g_df %>%
  ggplot(aes(x = Year,
             y = cum_sum_pct_chg,
             group = Glacier.Name,
             color = Glacier.Name))+
  geom_line(color = 'white',
            size = 2)+
  geom_point()+
  geom_line(size = 1.0)+
    # xlim(1966,2045)+
  geom_text_repel(data = g_df %>% filter(Year == 2015 &
                                     Glacier.Name %in% big_g$Glacier.Name), 
            aes(label = paste0(cum_sum_pct_chg, "%", " - ",Glacier.Name)) ,
            hjust = 'right',
            # hjust = 'right',
            size = 3.25,
            # nudge_x = 5,
            xlim = c(2025,2029),
            direction = "y",
            color = 'black',
            segment.size =0.1,
            segment.color = 'gray'
            # box.padding = 0.25
            
            ) +

  geom_text_repel(data = g_df %>% filter(Year == 1966 &
                                       Glacier.Name == "Agassiz Glacier"), 
              aes(label = paste0("100%")) ,
              hjust = 'left',
              # # hjust = 'right',
              size = 3.25,
              nudge_x = -2,
              # xlim = c(2025,2029),
              # direction = "y",
              color = 'black',
              # segment.size =0.1,
              # segment.color = 'gray'
              # box.padding = 0.25
              
              ) +
  
  MySpecial+
  
  #make axes margins larger
  # theme(plot.margin = margin(5,100,5,5))+
  #  Labelling as desired
  labs(
    title = "10 Largest Glaciers in Glacier National Park",
    subtitle = "Cumulative Percent Loss of Area",
    caption = "Data: Fagre et al., 2017"
  )

ggsave(filename = here("figures","Draft_Slope_plot.pdf"))

```

plot below with lines colored by decrease in area
```{r}
######
MySpecial <- list(
  
  scale_x_continuous(position = "top",
                   breaks = c(1966, 1998, 2005, 2015),
                   # expand = expansion(mult = c(0.01, 0.285)),
                   limits =  c(1966, 2030)),
  theme_bw(),
  # Format tweaks
  # tweak the legend
  theme(legend.position = 'bottom',
        legend.key.width = unit(1,'cm')),
        # legend.box = 'horizontal'),
  # theme(legend.key.width=unit(1,"cm")),
  # Remove the panel border
  theme(panel.border     = element_blank()),
  
  # Remove just about everything from the y axis
  theme(axis.title.y     = element_blank()),
  theme(axis.text.y      = element_blank()),
  theme(panel.grid.major.y = element_blank()),
  theme(panel.grid.minor.y = element_blank()),
  # Remove a few things from the x axis and increase font size
  theme(axis.title.x     = element_blank()),
  theme(panel.grid.major.x = element_blank()),
  theme(panel.grid.minor.x = element_blank()),
  theme(axis.text.x.top      = element_text(size=12)),
  # Remove x & y tick marks
  theme(axis.ticks       = element_blank()),
    # Format title & subtitle
  theme(plot.title       = element_text(size=14, face = "bold", hjust = 0.5)),
  theme(plot.subtitle    = element_text(hjust = 0.5))
  # theme(plot.margin = unit(c(0.1, 5, 0.1, 0.1), "cm"))
)

g_df %>%
  ggplot(aes(x = Year,
             y = cum_sum_pct_chg,
             group = Glacier.Name,
             color = ..y..))+
  geom_line(color = 'white',
            size = 2)+
  geom_line(size = 1.0)+
    # xlim(1966,2045)+
  geom_text_repel(data = g_df %>% filter(Year == 2015 &
                                     Glacier.Name %in% big_g$Glacier.Name), 
            aes(label = paste0(cum_sum_pct_chg, "%", " - ",Glacier.Name)) ,
            hjust = 0,
            # hjust = 'right',
            size = 3.25,
            # nudge_x = 5,
            xlim = c(2016, 2030),
            direction = "y",
            color = 'black',
            segment.size =0.2
            # box.padding = 0.25
            ) +
  # geom_label(aes(label = cum_sum_pct_chg), 
  #            size = 2.5, 
  #            label.padding = unit(0.05, "lines"), 
  #            label.size = 0.0) +
  MySpecial+
  
  #make axes margins larger
  # theme(plot.margin = margin(5,100,5,5))+
  #  Labelling as desired
  labs(
    title = "10 Largest Glaciers in Glacier National Park",
    subtitle = "Cumulative Percent Loss of Area",
    caption = "Data: Fagre et al., 2017",
    color = "% Loss"
  )



```